---
title: "Homework_3"
output: github_document
---
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggridges)
```


## Problem 1
```{r}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("instacart")  
```

This dataset contains `r nrow(instacart)` rows and ...columns.

Observations are the level of items in orders by user. There are some user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

How many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))

```

let's make a plot

```{r}
instacart %>% 
   count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```


Let's make a table!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


Apples vs ice cream

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```



## Problem 2

Import and tidy data
```{r}
accel_data = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate_at(vars(activity_1:activity_1400), as.numeric) %>% 
  pivot_longer(
    activity_1:activity_1400,
    names_to = "minute_of_day",
    values_to = "activity_count",
    names_prefix = "activity") %>% 
 mutate(weekday = case_when(
day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
day %in% c("Saturday","Sunday") ~ "weekend")) %>% 
  relocate("week", "weekday", "day", "day_id", "minute_of_day", "activity_count") %>% 
   mutate_at(vars("weekday","day"), as.factor) %>% 
  mutate_at(vars("day_id"), as.numeric)
    
accel_data
```

Total Activity variable + table

```{r}
agg_accel = 
accel_data %>% 
  group_by(day_id) %>% 
  summarise(across(starts_with('activity_1401'), sum)) 

rename(agg_accel, "activity_count" = "activity_1401") %>% 
  knitr::kable()
```

